---
title: 'SAX: Symbolic Aggregate Approximation'
author: "Ariel Lev, 27. February 2016"
output: 
  html_document:
    keep_md: true
    css: style.css
    pandoc_args: [
      "+RTS", "-K64m",
      "-RTS"
    ]    
---

```{r, echo=F, fig.width=12, fig.height=4, message=F, warning=F}

# loading packages

installed <- rownames(installed.packages())
required <- c("ggplot2", "knitr", "seewave", "devtools", "TSclust", "wavelets")

# filtering uninstalled packages
to_install <- required[required %in% installed == F]
if (length(to_install) > 0) {
  # installing
  print("following packages are missing and will be installed:")
  print(to_install)
  install.packages(to_install, repos = "http://cran.us.r-project.org")
}

git_hub_required <- c('Quandl')
to_install <- git_hub_required[git_hub_required %in% installed == F]
if (length(to_install) > 0) {
  # installing
  print("following packages are missing and will be installed:")
  print(to_install)
}

require(ggplot2)
require(knitr)
require(seewave)
require(Quandl)
require(rCharts)
require(TSclust)
require(stats)
require(dplyr)
require(wavelets)

#Quandl.api_key("cAFkxw5nDxxw1WJx8xB7")
# apple <- Quandl("WIKI/AAPL", trim_start="2015-01-01", trim_end="2016-01-01")  
```

### Getting and cleaning the data
Loading the Apple Inc (APPL) data set into R and performing an initial exploration.
```{r, echo=T, message=F}
apple <- read.table("apple.csv", stringsAsFactors = F)
apple$Date <- as.Date(apple$Date)
apple <- apple %>% arrange(Date)
kable(head(apple,10), caption = "Table 1.|Apple Inc (AAPL) Prices, Dividends, Splits and Trading Volume 2015")
```

### Dimension reduction using SAX
Translating the time series into its symbolic representation using SAX, and computing the frequency of each SAX symbol in this representation.
```{r, echo=T, message=F}

# amount of equal sized frames
W <- 50 

# The size of the alphabet
alpha <- 10

data <- apple$Open


# assumes that the distribution of the numeric series x follows a Gaussian distribution
# therefore normalizing
normalize <- (data - mean(data)) / sd(data)

# convert to SAX representation
data_paa <- PAA(normalize, W)
data_sax <- convert.to.SAX.symbol(data_paa, alpha)

# converting to letters
data_sax <- sapply(data_sax, function(x) LETTERS[x])
data_sax

# frequency
table(data_sax)
```

### Alternative dimensionality reduction method
```{r, echo=T, message=F}
# discrete wavelet transform
wt <- dwt(data,  n.levels=7, boundary="periodic", fast=FALSE)

# discrete Fourier transform
# data_dft <- FFT(data)
```

### Parallelize your implementation of SAX
```{r, echo=T, message=F}
# Can be implemented via a classic MapR job:
# 1. Map tasks iterate over sequental chunks of data for feature selection. 
# 2. A combiner averages features and yields PAA values (Piecewise Aggregate Approximation).
# 3. Reducer sort PAAs and apply SAX.
```

### A graphical interface
```{r, echo=T, message=F, results='asis', tidy=F, comment=NA}

## data wrangling for plot

# utility function to convert numeric vectors into data frames containing original dates
makeDF <- function(x, label, dates, type) {
  df <- as.data.frame(dates)
  names(df) <- c("Date")  
  df[label] <- x
  df$type <- type
  df
}

apple_df <- apple[,1:2]
apple_df$type <- "Stock"

# calculating PAA over an unnormalized data for the graphical representation
data_paa <- PAA(data, W)

data_paa <- rep(data_paa, each=length(data)/W)
data_sax <- rep(data_sax, each=length(data)/W)

# fixing length of p to align it with number of raw observations
data_paa <- append(data_paa, rep(data_paa[length(data_paa)], length(data) - length(data_paa)))
data_sax <- append(data_sax, rep(data_sax[length(data_sax)], length(data) - length(data_sax)))

paa_df <- makeDF(data_paa, "Open", apple$Date, "PAA")
#dwt_df <- makeDF(data_dwt, "Open", apple$Date, "dwt")
#dft_df <- makeDF(data_dft, "Open", apple$Date, "dft")

all <- rbind(apple_df, paa_df)

## plotting
p <- nPlot(Open ~ Date, type = "lineChart", data = all, group = "type")
p$xAxis(
    tickFormat =   "#!
      function(d) {return d3.time.format('%B')(new Date(d*1000*3600*24));}
    !#"
  )

p$chart(tooltipContent = "#! function(key, x, y){
        return '<h3>' + key + '</h3>' + 
        '<p>' + y + ' in ' + x + '</p>'
        } !#")

p$addParams(width = 1000, pointSize = 0, lineWidth = 1)
p$xAxis(axisLabel = "Month")
p$yAxis(axisLabel = "Rate")
#p$chart(stacked = TRUE)
#p$chart(reduceXTicks = FALSE)
#p$chart(color = c('#db1847', '#169998'))
p$print('iframesrc', include_assets = TRUE)

```